### 2007-01-10 Documentation/part2.tex
*** empty log message ***

\index{aedl}{{\tt rule}|(}
The set of rules\index{norm}{Rule} for an experiment is the final 
link between the event rate computation and the statistical analysis. 
The information in the rules
specifies how the $\chi^2$ is computed based upon the raw event rates 
given by the channels and possible systematical errors. 
Therefore a rule has two parts: The first part describes how signal and 
background events are composed out of the channels, and the second part
specifies which systematical errors are considered, as well as their values.
%
For a rule, the splitting
in signal and background is useful for the treatment of systematics, as we will se later. Each rule will lead to a $\Delta \chi^2$-value,
which means that all $\Delta \chi^2$'s of the different rules will be added
for the whole experiment. Within each rule, the event rates are added, and
the systematics is considered to be independent of the other rules.
Thus, it is convenient to combine the above defined channels for different
oscillation patterns and interaction types into one logical construction,
which is the rule. For example, a superbeam usually has two rules: One for
the $\nu_e$-appearance rates, and one for the $\nu_\mu$-disappearance rates.
 In each case, contributions of several interaction types, as well as from
 the $\nu_e$-contamination of the beam will lead to a number of contributing signal and background event channels.

@@

For the analysis of the systematical errors, the so called 
 ``pull method'' is used~\cite{Fogli:2002pt}\footnote{In fact the pull 
method was employed already in \Ref~\cite{Huber:2002mx} before 
\Ref~\cite{Fogli:2002pt} appeared.}. 
For the pull method, $k$ systematical errors are included by introducing 
$k$ additional variables $\zeta_k$, which are the 
so-called ``nuisance parameters''. 
The nuisance parameters describe the dependence 
of the event rates on the various sources of systematical errors, 
such as an error on the total 
normalization is included by multiplying the expected number of events in 
each bin by a factor $(1+\zeta_1)$. The variation of $\zeta_1$ is in the fit 
constrained by adding a penalty $p_1$ to the $\chi^2$-function. In case 
of a  Gau\ss ian distributed systematical error, this penalty is 
given by
\begin{equation}
\label{eq:penalty}
p_i=\frac{(\zeta_i-\zeta_i^0)^2}{\sigma_{\zeta_i}^2}\,,
\end{equation}
where $\zeta_i^0$ denotes the mean and $\sigma_{\zeta_i}$ the standard 
deviation of the corresponding nuisance parameter. We further on also refer
to the mean as the ``central value'', and to the standard deviation as
the ``error''. The latter corresponds to the actual systematical uncertainty. 
The resulting $\chi^2$ is then minimized with respect to all nuisance 
parameters $\zeta_i$, which leads to $\chi^2_\mathrm{pull}$
\begin{equation}
\chi^2_\mathrm{pull}(\boldsymbol{\lambda}):=\min_{\{\zeta_i\} } \,\, \left( 
\chi^2(\boldsymbol{\lambda},
\zeta_1, \ldots, \zeta_k)+ \sum_{j=1}^{k} p_j(\zeta_j)\right)\,.
\end{equation}
Here $\boldsymbol{\lambda}$ refers to the oscillation parameters 
including the matter density
$\rho$. One advantage of the pull method is that whenever the number $N$ of 
data points is much larger than $k$, it is numerically easier to compute 
$\chi^2_\mathrm{pull}$ than to invert the $N\times N$ covariance matrix. For
the experiments considered here, $N$ is typically $20$ and $k\sim 4$, 
which means that the pull method is numerically much faster. Moreover,
 it is more flexible and  allows the inclusion of systematical errors 
also for a  Poissonian $\chi^2$-function.
In \Ref~\cite{Fogli:2002pt}, it has been demonstrated that the pull method 
and the covariance based approach are equivalent for a Gau\ss ian and 
linear model. In general,
there is a separate $(\chi^2_\mathrm{pull})^r$ for each rule $r$, 
\ie , pair of signal and background spectra, with a separate set of 
nuisance parameters $\zeta_i^\alpha$. Thus, $\chi^2_\mathrm{pull}$
is the sum of all individual  $(\chi^2_\mathrm{pull})^r$'s.
By the minimization, the dependence on the $k$ nuisance parameters has been eliminated  from $\chi^2_\mathrm{pull}$.

Now we can introduce the different systematical errors. 
The two most important and
most easily parameterized systematical errors are the normalization 
and an energy calibration errors. These errors are assumed to be independent between the signal events and the background events, which means that
this systematics treatment defines the grouping into signal or background.
The implementation of the normalization error
is straightforward:
\begin{equation}
s_i(a):=a\cdot s_i
\end{equation} 
with an analogous definition for the background events. Here, $a$ is the ``nuisance'' parameter, which will be minimized over later.

For the parameterization of an energy calibration error, two possibilities
are implemented. The first one (method ``T'') is somewhat simpler, 
whereas the second one (method ``C'')
is more accurate, but it requires a careful choice of parameters. 
The first option (method ``T'') is (JK: IS THAT CORRECT? AND ALSO THE NEXT EQUATION? (THERE HAVE BEEN
SOME CHANGES) [WW])
\begin{equation}
s_i(a,b) \equiv s_i(a)+b\cdot s_i\, E'_i/(E'_\mathrm{max}-E'_\mathrm{min}),
\end{equation}
where $E'_{\mathrm{min}}$ and $E'_{\mathrm{max}}$ correspond to {\tt \$emin}
and {\tt \$emax}, and $E'_i$ is the mean (reconstructed) 
energy of the $i$th bin. It is often refered to as a ``tilt'' of the
spectrum, since it describes a linear distortion 
of the event rate spectrum. Note that this definition of a tilt makes
only sense in combination with a large enough normalization error, since
the tilt also affects the normalization.
%
The second option (method ``C'') is closer to an actual energy
calibration error, which means that one should test this option whenever
one suspects a large impact of this systematical error.
It is based upon replacing the events in the $i$th bin by the ones at
the energy $(1+b)\cdot E'_i$. If this target energy does not exactly hit
a (discrete) bin energy $E_k$, linear interpolation is used. We use the following approximation:
\begin{eqnarray}
s_i(a,b)&=& (1+b)\cdot \left[ \left( s_{k+1}(a)-s_k(a) \right)\cdot (\delta-k) +s_k(a) \right] \,,\\
\delta&=&b\cdot(i+t_0+ 1/2)+i\,,\nonumber\\
k&=& \mathrm{div}(\delta,1)\,,\nonumber\\
t_0&=&E'_\mathrm{min}/\Delta E_0\,.\nonumber
\end{eqnarray}
Here $\Delta E_0$ is the bin width ({\tt \$emax}-{\tt \$emin})/{\tt \$bins}
and ``div'' refers to the integer part of the division. Note that the
factor $(1+b)$ in the first equation comes from a renormalization of
the bin width, since also the bin width is altered by the replacement of
the energies. Furthermore, special care has to be payed to the limits
 $k<1$ or $k+1>N_\mathrm{bins}$, since there $s_k$ or $s_{k+1}$ may not have been calculated. By default, it is assumed that $s_k$ is
zero in those cases. However, if the event rates are still large at
the limits, errors will be introduced leading to a wrong
estimate of the impact of the calibration error. In this case,
one should truncate the analysis range by a few bins at the boundaries
and therefore ensure in this way that only those $s_i$ are used whose index
$k$ is within the range $0,\ldots, N_\mathrm{bins}-1$ (\cf, \figu{energies}). 
Thus, it is possible to constrain the analysis energy range 
with each rule to an energy window\index{norm}{Energy!window}:
\index{aedl}{{\tt rule}!energy window@{\tt "@energy\_window}}
\begin{quote}
{\tt 
\tb @energy\_window = 4.0 : 50.0 
}
\end{quote}
The default energy window is given between the minimal and maximal reconstructed energy. To be on the save side, reduce analysis window
compared to the bin range on each side by about three times the
energy calibration error.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[t!]
\begin{center}
\begin{tabular}[h]{|lccccccl|}
\hline
Systematics function &$a$&$b$&$c$&$d$&Tilt/Calibr.& Dim. & Remarks\\
\hline
\hline
\multicolumn{8}{|l|}{{\bf Standard systematics:}} \\
{\tt chiSpectrumTilt} &+&+&+&+&T&0&Systematics with tilt\\
{\tt chiNoSysSpectrum} &-&-&-&-&-&2&No systematics\\
{\tt chiTotalRatesTilt} &+&+&+&+&T&4&Total rates\\
{\tt chiSpectrumOnly} & $\infty$ &-& $\infty$ &-&-&7&Spectrum only\\
{\tt chiNoSysTotalRates} &-&-&-&-&-&8&Total rates, no syst.\\
{\tt chiSpectrumCalib} &+&+&+&+&C&9&ystematics with calibr.\\
\hline\multicolumn{8}{|l|}{{\bf User-defined systematics:}} \\
{\tt chiZero} & - & - & - & - & - & n/a & Passive rule\\
 & & & & & & & ($\chi^2$ returns zero) \\
Any other name & \multicolumn{5}{c}{User-defined behavior} & n/a & User-defined syst. \\
\hline
\end{tabular}
\mycaption[Table of error dimensions]{\label{tab:error_dim}
Possible systematics $\chi^2$ functions  in \GLOBES\ and their meaning. 
If a parameter is designated with $+$, it will
be marginalized over, and therefore the corresponding error needs to 
have a non-zero value. If the cases with ``total rates'' in the remarks, 
the summation over the bins is performed \emph{before} computing 
the $\chi^2$, \ie, no spectral information is used. 
The function {\tt chiSpectrumOnly} leaves the 
normalization free ($\sigma_a=\sigma_c=\infty$), and therefore 
only the spectral information is used. As a consequence,
the settings for the normalization error will be ignored (designated 
with the symbol $\infty$). In addition, the corresponding error
dimension from earlier versions of \GLOBES\ is shown in the column ``Dim.''.
 }
\end{center}
\index{aedl}{{\tt rule}!errordim@{\tt "@errordim}} 
\index{norm}{Systematics function} 
\end{table} 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Eventually, the total event rate $x_i$ in a bin $i$ is given by
\begin{equation}
x_i(a,b,c,d)=s_i(a,b)+b_i(c,d) \, ,
\end{equation}
and is thus a function of four parameters. 
The four parameters $a,b,c,d$ have been introduced in order to describe
systematical uncertainties and are the nuisance parameters.
Each of the four parameters has a central value and systematical error.
The central values for all of
the four parameters have to be always defined. They are called 
signal normalization ($a$), signal tilt/calibration ($b$), 
background  normalization ($c$) and
background tilt/calibration ($d$). The default values are
\begin{equation}
a=1\,,\quad b=0\,,\quad c={\tt not \, \, assigned}\,,\quad d=0\,.
\end{equation}
Thus, for the background normalization $c$, the value has to be specified 
in \emph{either} case. The values for the normalization and the 
values of tilt/calibration are always regarded
as a pairs, \ie, they are given in the form {\tt normalization : tilt}. The errors are treated in the same way. For example, we have
\index{aedl}{{\tt rule}!signal error@{\tt "@signalerror}}
\index{aedl}{{\tt rule}!background error@{\tt "@backgrounderror}}
\index{aedl}{{\tt rule}!background center@{\tt "@backgroundcenter}}
\begin{quote}
{\tt
\tb @signalerror =       0.001  :       0.01\\
\tb @backgroundcenter =  0.1 :       0.0\\
\tb @backgrounderror =   0.001 :       0.01
}
\end{quote}
There is no {\tt @signalcenter} in this definition, 
since by default the central value for the
signal normalization is $1$ and the central value for the tilt/calibration 
is $0$.

The user has the possibility to choose the set $\{\zeta_i\}$ of nuisance 
parameters which are minimized over. This choice is specified with the 
error dimension variable\index{norm}{Error dimension}, and the different 
possibilities are shown in
\Tab~\ref{tab:error_dim}. Since the error dimension defines the treatement
of systematics it is useful to have define a matched pair of error dimensions
for each rule, where on value describes how the event rate is computed
for \emph{no} systematics and the other one is with systematics 
(see also~\Sec~\ref{sec:systematics}).
The error dimensions for the case of no systematics is set for each rule 
by the value of {\tt @errordim\_sys\_off}\index{aedl}{{\tt rule}!errordim 
sys off@{\tt "@errordim\_sys\_off}}, whereas the error dimension
for systematics on is given by {\tt @errordim\_sys\_on}
\index{aedl}{{\tt rule}!errordim sys on@{\tt "@errordim\_sys\_on}}. Thus,
 for example,
the complete error dimension definition could look like  
\begin{quote}
{\tt 
\tb @errordim\_sys\_off = 2\\
\tb @errordim\_sys\_on = 0
}
\end{quote}
If none of the built-in algorithms listed in \Tab~\ref{tab:error_dim} is
suitable for a certain setup, the user can override them by setting
{\tt @errordim\_sys\_on} to 20 and using the \GLB{SetUserChi}
function to specify a specialized function for the calculation
of $\chi^2$. This function will then be called by the \GLOBES\ minimizers.
For details, see \Sec~\ref{sec:userchi}.

Eventually, a rule looks like
\begin{quote}
{\tt rule(\#rule\_1)<\\
\tb @signal = 0.5 @ \#channel\_1\\
\tb @background = 0.001 @ \#channel\_2 :  0.005 @ \#channel\_3\\
\tb @signalerror =       0.001  :       0.01\\
\tb @backgroundcenter =  0.1 :       0.0\\
\tb @backgrounderror =   0.001 :       0.01\\
\tb @errordim\_sys\_off = 2\\
\tb @errordim\_sys\_on = 0\\
\tb @energy\_window = 4.0 : 50.0\\ 
>}
\end{quote}
\index{aedl}{{\tt rule}|)}

@@

\index{norm}{globes@{\tt globes}!oscillation parameters}
Calling {\tt globes} without any options and with a {\tt .glb}-file as argument produces an event summary at rule level. In this case,  
the full experiment description in the file is taken into account, \ie, all efficiencies, backgrounds, and energy resolution effects. 
Thus, the returned event rates are the ones which will be actually 
used to compute the $\chi^2$ later. By default, the oscillation parameters used to calculate the transition probability are
\begin{eqnarray}
\label{eq:globes_params}
\sin^22\theta_{12}=0.8&\quad&\Delta m^2_{21}=7\cdot10^{-5}
\,\mathrm{eV}^2\,,\nonumber\\
\sin^22\theta_{23}=1.0&\quad&\Delta m^2_{31}=3\cdot10^{-3}
\,\mathrm{eV}^2\,,\nonumber\\
\delta=0&\quad&\sin^22\theta_{13}=0.1\,.\
\end{eqnarray}
Of course, it is possible to change these default values either by using the
option {\tt -p} on a call by call basis, or by setting the environment variable
{\tt GLB\_CENTRAL\_VALUES}:
%
\index{norm}{GLB CENTRAL VALUES@{\tt GLB\_CENTRAL\_VALUES}} 
\index{norm}{Environment variables!GLB CENTRAL VALUES@{\tt GLB\_CENTRAL\_VALUES}}
%
\begin{quote}
{\tt
globes -p'0.55,0,0.785,0,0.0008,0.0025'\\
globes \verb+--+parameters='0.55,0,0.785,0,0.0008,0.0025'
}
\end{quote}
For example, {\tt GLB\_CENTRAL\_VALUES} can be defined
within the shell session or in the shell profile:
\begin{quote}
{\tt
export GLB\_CENTRAL\_VALUES='0.55,0,0.785,0,0.0008,0.0025'
}
\end{quote}
Furthermore it is possible to switch off oscillations with the {\tt -N} option and to switch them on again with {\tt -O} (the default). The effect of {\tt -N} is the same as to use {\tt NOSC\_} in all oscillation channels.
Thi feature is useful if one wants to normalize an expriment flux if
the number of un-oscillated events is given.

@@

It is also possible for rules to switch off one detector effect after 
the other -- with the limitation that rules only make sense
after the energy resolution function has been applied to each channel.
Therefore, it is \emph{not} possible to use {\tt -b} together with {\tt -r}, or to switch off any pre-smearing efficiencies or backgrounds. 
One can, however, switch off the post-smearing efficiencies ({\tt -f}) 
and the post-smearing backgrounds ({\tt -g}) for each channel. Since
the definition of a rule also contains so-called ``coefficients'', it is
possible to switch them off with {\tt -i}. This options also deactivates
any setting of {\tt @backgroundcenter}.

>>>>

\index{aedl}{{\tt rule}|(}
The set of rules\index{norm}{Rule} for an experiment is the final 
link between the event rate computation and the statistical analysis. 
The information in the rules
specifies how the $\chi^2$ is computed based upon the raw event rates 
given by the channels and possible systematical errors. 
Therefore a rule has two parts: The first part describes how signal and 
background events are composed out of the channels, and the second part
specifies which systematical errors are considered, as well as their values.
%
For a rule, the splitting
in signal and background is useful for the treatment of systematics, as we will se later. Each rule will lead to a $\Delta \chi^2$-value,
which means that all $\Delta \chi^2$'s of the different rules will be added
for the whole experiment. Within each rule, the event rates are added, and
the systematics is considered to be independent of the other rules (unless user-defined
systematics specifies a dependence).
Thus, it is convenient to combine the above defined channels for different
oscillation patterns and interaction types into one logical construction,
which is the rule. For example, a superbeam usually has two rules: One for
the $\nu_e$-appearance rates, and one for the $\nu_\mu$-disappearance rates.
 In each case, contributions of several interaction types, as well as from
 the $\nu_e$-contamination of the beam will lead to a number of contributing signal and background event channels.

@@

For the analysis of the systematical errors, the so called 
 ``pull method'' is used~\cite{Fogli:2002pt}\footnote{In fact the pull 
method was employed already in \Ref~\cite{Huber:2002mx} before 
\Ref~\cite{Fogli:2002pt} appeared.}. 
For the pull method, $k$ systematical errors are included by introducing 
$k$ additional variables $\zeta_k$, which are the 
so-called ``nuisance parameters''. 
The nuisance parameters describe the dependence 
of the event rates on the various sources of systematical errors, 
such as an error on the total 
normalization is included by multiplying the expected number of events in 
each bin by a factor $(1+\zeta_1)$. The variation of $\zeta_1$ is in the fit 
constrained by adding a penalty $p_1$ to the $\chi^2$-function. In case 
of a  Gau\ss ian distributed systematical error, this penalty is 
given by
\begin{equation}
\label{eq:penalty}
p_i=\frac{\zeta_i^2}{\sigma_{\zeta_i}^2}\,,
\end{equation}
where $\sigma_{\zeta_i}$ is the standard 
deviation of the corresponding nuisance parameter. We further on also refer
to the  standard deviation as
the ``error'', it corresponds to the actual systematical uncertainty.
Note that the central values of all penalties are zero in \GLOBES\ 3.0 and higher. 
The resulting $\chi^2$ is then minimized with respect to all nuisance 
parameters $\zeta_i$, which leads to $\chi^2_\mathrm{pull}$
\begin{equation}
\chi^2_\mathrm{pull}(\boldsymbol{\lambda}):=\min_{\{\zeta_i\} } \,\, \left( 
\chi^2(\boldsymbol{\lambda},
\zeta_1, \ldots, \zeta_k)+ \sum_{j=1}^{k} p_j(\zeta_j)\right)\,.
\end{equation}
Here $\boldsymbol{\lambda}$ refers to the oscillation parameters 
including the matter density
$\rho$. One advantage of the pull method is that whenever the number $N$ of 
data points is much larger than $k$, it is numerically easier to compute 
$\chi^2_\mathrm{pull}$ than to invert the $N\times N$ covariance matrix. For
the experiments considered here, $N$ is typically $20$ and $k\sim 4$, 
which means that the pull method is numerically much faster. Moreover,
 it is more flexible and  allows the inclusion of systematical errors 
also for a Poissonian $\chi^2$-function.
In \Ref~\cite{Fogli:2002pt}, it has been demonstrated that the pull method 
and the covariance based approach are equivalent for a Gau\ss ian and 
linear model. In general,
there is a separate $(\chi^2_\mathrm{pull})^r$ for each rule $r$, 
\ie , pair of signal and background spectra, with a separate set of 
nuisance parameters $\zeta_i^\alpha$. Thus, $\chi^2_\mathrm{pull}$
is the sum of all individual  $(\chi^2_\mathrm{pull})^r$'s.
By the minimization, the dependence on the $k$ nuisance parameters has been eliminated  from $\chi^2_\mathrm{pull}$.

Now we can introduce the different systematical errors. 
The two most important and
most easily parameterized systematical errors are the normalization 
and an energy calibration errors. These errors are assumed to be independent between the signal events and the background events, which means that
this systematics treatment defines the grouping into signal or background.
The implementation of the normalization error
is straightforward:
\begin{equation}
s_i(a):=(1+a)\cdot s_i
\end{equation} 
with an analogous definition for the background events. Here, $a$ is the ``nuisance'' parameter, which will be minimized over later.

For the parameterization of an energy calibration error, two possibilities
are implemented. The first one (method ``T'') is somewhat simpler, 
whereas the second one (method ``C'')
is more accurate, but it requires a careful choice of parameters. 
The first option (method ``T'') is (JK: IS THAT CORRECT? AND ALSO THE NEXT EQUATION? (THERE HAVE BEEN
SOME CHANGES) [WW])
\begin{equation}
s_i(a,b) \equiv s_i(a)+b\cdot s_i\, E'_i/(E'_\mathrm{max}-E'_\mathrm{min}),
\end{equation}
where $E'_{\mathrm{min}}$ and $E'_{\mathrm{max}}$ correspond to {\tt \$emin}
and {\tt \$emax}, and $E'_i$ is the mean (reconstructed) 
energy of the $i$th bin. It is often refered to as a ``tilt'' of the
spectrum, since it describes a linear distortion 
of the event rate spectrum. Note that this definition of a tilt makes
only sense in combination with a large enough normalization error, since
the tilt may also affect the normalization. 
%
The second option (method ``C'') is closer to an actual energy
calibration error, which means that one should test this option whenever
one suspects a large impact of this systematical error.
It is based upon replacing the events in the $i$th bin by the ones at
the energy $(1+b)\cdot E'_i$. If this target energy does not exactly hit
a (discrete) bin energy $E_k$, linear interpolation is used. We use the following approximation:
\begin{eqnarray}
s_i(a,b)&=& (1+b)\cdot \left[ \left( s_{k+1}(a)-s_k(a) \right)\cdot (\delta-k) +s_k(a) \right] \,,\\
\delta&=&b\cdot(i+t_0+ 1/2)+i\,,\nonumber\\
k&=& \mathrm{div}(\delta,1)\,,\nonumber\\
t_0&=&E'_\mathrm{min}/\Delta E_0\,.\nonumber
\end{eqnarray}
Here $\Delta E_0$ is the bin width ({\tt \$emax}-{\tt \$emin})/{\tt \$bins}
and ``div'' refers to the integer part of the division. Note that the
factor $(1+b)$ in the first equation comes from a renormalization of
the bin width, since also the bin width is altered by the replacement of
the energies. Furthermore, special care has to be payed to the limits
 $k<1$ or $k+1>N_\mathrm{bins}$, since there $s_k$ or $s_{k+1}$ may not have been calculated. By default, it is assumed that $s_k$ is
zero in those cases. However, if the event rates are still large at
the limits, errors will be introduced leading to a wrong
estimate of the impact of the calibration error. In this case,
one should truncate the analysis range by a few bins at the boundaries
and therefore ensure in this way that only those $s_i$ are used whose index
$k$ is within the range $0,\ldots, N_\mathrm{bins}-1$ (\cf, \figu{energies}). 
Thus, it is possible to constrain the analysis energy range 
with each rule to an energy window\index{norm}{Energy!window}:
\index{aedl}{{\tt rule}!energy window@{\tt "@energy\_window}}
\begin{quote}
{\tt 
\tb @energy\_window = 4.0 : 50.0 
}
\end{quote}
The default energy window is given between the minimal and maximal reconstructed energy. To be on the save side, reduce analysis window
compared to the bin range on each side by about three times the
energy calibration error.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[t!]
\begin{center}
\begin{tabular}[h]{|lccccccl|}
\hline
Systematics function &$a$&$b$&$c$&$d$&Tilt/Calibr.& Dim. & Remarks\\
\hline
\hline
\multicolumn{8}{|l|}{{\bf Standard systematics:}} \\
{\tt chiSpectrumTilt} &+&+&+&+&T&0&Systematics with tilt\\
{\tt chiNoSysSpectrum} &-&-&-&-&-&2&No systematics, but\\
 & & & & & & & spectral information \\
{\tt chiTotalRatesTilt} &+&+&+&+&T&4&Total rates\\
{\tt chiSpectrumOnly} & $\infty$ &-& $\infty$ &-&-&7&Spectrum only\\
{\tt chiNoSysTotalRates} &-&-&-&-&-&8&Total rates, no syst.\\
{\tt chiSpectrumCalib} &+&+&+&+&C&9&Systematics with calibr.\\
\hline\multicolumn{8}{|l|}{{\bf User-defined systematics:}} \\
{\tt chiZero} & - & - & - & - & - & n/a & Passive rule\\
 & & & & & & & ($\chi^2$ returns zero) \\
Any other name & \multicolumn{5}{c}{User-defined behavior} & n/a & User-defined syst. \\
\hline
\end{tabular}
\mycaption[Table of error dimensions]{\label{tab:error_dim}
Possible systematics $\chi^2$ functions  in \GLOBES\ and their meaning. 
If a parameter is designated with $+$, it will
be marginalized over, and therefore the corresponding error needs to 
have a non-zero value. If the cases with ``total rates'' in the remarks, 
the summation over the bins is performed \emph{before} computing 
the $\chi^2$, \ie, no spectral information is used. 
The function {\tt chiSpectrumOnly} leaves the 
normalization free ($\sigma_a=\sigma_c=\infty$), and therefore 
only the spectral information is used. As a consequence,
the settings for the normalization error will be ignored (designated 
with the symbol $\infty$). In addition, the corresponding error
dimension from earlier versions of \GLOBES\ is shown in the column ``Dim.''.
 }
\end{center}
%\index{aedl}{{\tt rule}!errordim@{\tt "@errordim}} 
\index{norm}{Systematics function} 
\end{table} 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Eventually, the total event rate $x_i$ in a bin $i$ is given by
\begin{equation}
x_i(a,b,c,d)=s_i(a,b)+b_i(c,d) \, ,
\end{equation}
and is thus a function of four parameters. 
The four parameters $a,b,c,d$ have been introduced in order to describe
systematical uncertainties and are the nuisance parameters.
Each of the four parameters has a corresponding systematical error.
% The central values for all of
% the four parameters have to be always defined for built-in systematics. 
They are called  signal normalization ($a$), signal tilt/calibration ($b$), 
background  normalization ($c$) and
background tilt/calibration ($d$). Their default (central) values are zero.
% \begin{equation}
% a=1\,,\quad b=0\,,\quad c={\tt not \, \, assigned}\,,\quad d=0\,.
% \end{equation}
% Thus, for the background normalization $c$, the value has to be specified 
% in \emph{either} case. 
The errors for the normalization and the 
values of tilt/calibration are always regarded
as a pairs, \ie, they are given in the form {\tt normalization : tilt}. For example, we have\footnote{The old parameter {\tt @backgroundcenter} should not be used anymore. A background normalization center of 1.0 will be
interpreted as zero central value. IS THAT CORRECT [WW]?}
\index{aedl}{{\tt rule}!signal error@{\tt "@signalerror}}
\index{aedl}{{\tt rule}!background error@{\tt "@backgrounderror}}
%\index{aedl}{{\tt rule}!background center@{\tt "@backgroundcenter}}
\begin{quote}
{\tt
\tb @signalerror =       0.001  :       0.01\\
%\tb @backgroundcenter =  0.1 :       0.0\\
\tb @backgrounderror =   0.001 :       0.01
}
\end{quote}
% There is no {\tt @signalcenter} in this definition, 
% since by default the central value for the
% signal normalization is $1$ and the central value for the tilt/calibration 
% is $0$.  
%
The user has the possibility to choose the set $\{\zeta_i\}$ of nuisance 
parameters which are minimized over. This choice is specified with the 
systematics functions {\tt sys\_on\_function} and {\tt sys\_off\_function}
corresponding two systematics modes ``systematics on'' and ``systematics off''.\footnote{In earlier
versions before \GLOBES\ 3.0 the ``error dimension'' was used. The corresponding
parameters {\tt errordim\_sys\_on} and {\tt errordim\_sys\_off} are still supported
but should not be used anymore.}
\index{aedl}{{\tt rule}!sys on function@{\tt "@sys\_on\_function}}
\index{aedl}{{\tt rule}!sys off function@{\tt "@sys\_off\_function}}
\index{norm}{Systematics function}
The different  possibilities are shown in
\Tab~\ref{tab:error_dim}. Since the dual systematics modes define the experiment
behavior for systematics on and off, it is useful to have a matching pair of 
systematics functions for each rule 
(see also~\Sec~\ref{sec:systematics}).. The signal and background errors specified
by {\tt @signalerror} and  {\tt @backgrounderror} will then be used, if applicable.
For example, one may define 
\begin{quote}
{\tt
\tb @signalerror =       0.001  :       0.01\\
\tb @backgrounderror =   0.001 :       0.01 \\
\tb @sys\_on\_function = "chiSpectrumTilt"  \\
\tb @sys\_off\_function = "chiNoSysSpectrum"  
}
\end{quote}

For user-defined systematics \index{norm}{User-defined systematics}
(see, \Sec~\ref{sec:userchi}), one can use arbitrary names
for the systematics functions which are not pre-defined. In this case,
one would specify the systematical errors in lists, such as
\index{aedl}{{\tt rule}!sys on errors@{\tt "@sys\_on\_errors}}
\index{aedl}{{\tt rule}!sys off errors@{\tt "@sys\_off\_errors}}
\begin{quote}
{\tt
\tb @sys\_on\_function = "chiMySystematics"  \\
\tb @sys\_on\_errors = \{ 0.2, 0.3, 0.5 \}~// Uses three syst. errors \\
\tb @sys\_off\_function = "chiNoSysSpectrum"  \\
\tb @sys\_off\_errors = \{ \}
}
\end{quote}
WHICH SYSTEMATICAL ERRORS WILL A FUNCTION USE IF THERE IS A MIXED DECLARATION,
SUCH AS WITH SYS\_ON\_ERRORS IN SYS ON AND SIGNALERROR/BACKGROUNDERROR in SYS OFF?
DOES THAT WORK? [WW]
Note that here the systematics on and systematics off models can be used 
as fully functional different systematics modes with different systematical
errors. The interpretation of the systematical errors in {\tt @sys\_on\_errors} and
{\tt @sys\_off\_errors} is left to the application software code and the user. 
In addition, the application software has to register the user-defined systematics $\chi^2$
function with \GLB{DefineChiFunction} and uniquely identify the name
given by {\tt @sys\_on\_function} or {\tt @sys\_off\_function} to avoid
confusing different systematics routines. In addition, it is possible to define
a rule with passive systematics using {\tt glbChiZero}. Such passive systematics
is not directly used or called, but the corresponding rate vectors will be calculated
and provided for indirect access. For example, a reactor experiment with correlated
systematics between near and far detectors may define the user-defined systematics
{\tt chiReactor} for the far detector and use {\tt chiZero} for the near detector.
The routine assigned to the far detector will then perform the $\chi^2$ calculation,
which, of course, also involves the rates of the near detector. However, defining
{\tt chiReactor} in both the near and far detectors would result in a double call
of the $\chi^2$ function, \ie, the program will be a factor two slower and the
resulting $\chi^2$ will be a factor two too large.

Eventually, a complete rule may look like this:
\begin{quote}
{\tt rule(\#rule\_1)<\\
\tb @signal = 0.5 @ \#channel\_1\\
\tb @background = 0.001 @ \#channel\_2 :  0.005 @ \#channel\_3\\
\tb @signalerror =       0.001  :       0.01\\
\tb @backgrounderror =   0.001 :       0.01\\
\tb @sys\_on\_function = "chiSpectrumTilt"  \\
\tb @sys\_off\_function = "chiNoSysSpectrum"  \\
\tb @energy\_window = 4.0 : 50.0
>}
\end{quote}
\index{aedl}{{\tt rule}|)}

@@

\index{norm}{globes@{\tt globes}!oscillation parameters}
Calling {\tt globes} without any options and with a {\tt .glb}-file as argument produces an event summary at rule level. In this case,  
the full experiment description in the file is taken into account, \ie, all efficiencies, backgrounds, and energy resolution effects. 
Thus, the returned event rates are the ones which will be actually 
used to compute the $\chi^2$ later. By default, the oscillation parameters used to calculate the transition probability are
\begin{eqnarray}
\label{eq:globes_params}
\sin^22\theta_{12}=0.8&\quad&\Delta m^2_{21}=7\cdot10^{-5}
\,\mathrm{eV}^2\,,\nonumber\\
\sin^22\theta_{23}=1.0&\quad&\Delta m^2_{31}=3\cdot10^{-3}
\,\mathrm{eV}^2\,,\nonumber\\
\delta=0&\quad&\sin^22\theta_{13}=0.1\,.\
\end{eqnarray}
Of course, it is possible to change these default values either by using the
option {\tt -p} on a call by call basis, or by setting the environment variable
{\tt GLB\_CENTRAL\_VALUES}:
%
\index{norm}{GLB CENTRAL VALUES@{\tt GLB\_CENTRAL\_VALUES}} 
\index{norm}{Environment variables!GLB CENTRAL VALUES@{\tt GLB\_CENTRAL\_VALUES}}
%
\begin{quote}
{\tt
globes -p'0.55,0,0.785,0,0.0008,0.0025'\\
globes \verb+--+parameters='0.55,0,0.785,0,0.0008,0.0025'
}
\end{quote}
For example, {\tt GLB\_CENTRAL\_VALUES} can be defined
within the shell session or in the shell profile:
\begin{quote}
{\tt
export GLB\_CENTRAL\_VALUES='0.55,0,0.785,0,0.0008,0.0025'
}
\end{quote}
Note that in the case of additional non-standard parameters, these will just to 
be included at the end of these lists.
Furthermore it is possible to switch off oscillations with the {\tt -N} option and to switch them on again with {\tt -O} (the default). The effect of {\tt -N} is the same as to use {\tt NOSC\_} in all oscillation channels.
Thi feature is useful if one wants to normalize an expriment flux if
the number of un-oscillated events is given.

@@

It is also possible for rules to switch off one detector effect after 
the other -- with the limitation that rules only make sense
after the energy resolution function has been applied to each channel.
Therefore, it is \emph{not} possible to use {\tt -b} together with {\tt -r}, or to switch off any pre-smearing efficiencies or backgrounds. 
One can, however, switch off the post-smearing efficiencies ({\tt -f}) 
and the post-smearing backgrounds ({\tt -g}) for each channel. Since
the definition of a rule also contains so-called ``coefficients'', it is
possible to switch them off with {\tt -i}. 
% This options also deactivates
% any setting of {\tt @backgroundcenter}.
